---
title: "52414: Lab 2"
author: "Adi Ziv"
date: "May 19, 2020"
output: html_document
---

Required Libraries:
```{r, echo=FALSE}
library(ggplot2)
library(dplyr)
library(corrplot)
library(scales)   # needed for formatting y-axis labels to non-scientific type
library(radarchart)
library(tidyr)
library(tidyverse)
library(reshape2) # melt
library(ggthemes)
library(rworldmap) # world map
library(modelr)
library(radarchart) #Spider chart
library(RColorBrewer) #to make it colorful map
```

<br/><br/>


Loading the data:
```{r}
fifa_players <- data.frame(read.csv(url("https://raw.githubusercontent.com/DataScienceHU/DataAnalysisR_2020/master/data/fifa_data.csv")))
# Pre-processing: 
for (i in c(3,6,7,10:71)) {
  fifa_players[,i]<-as.numeric((fifa_players[,i]))
}
fifa<-na.omit(fifa_players)
fifa_players<-fifa
fifa_players_info<-fifa[,c(1:11)] # players general info 
fifa_players_attribures<-fifa[,c(1,12:45, 6)] # players different skills. Add overall
fifa_players_positions<-fifa[,c(1,46:72,6)] # players ability at different positions . Add overall 
```

```{r}
ggplot(fifa_players_info, aes(x=Age)) + geom_histogram(binwidth=1, color="black", fill="white")
```

```{r}
ggplot(fifa_players_info, aes(x=League, y=Overall, fill=League)) + 
  geom_boxplot(outlier.colour="black", outlier.shape=16, outlier.size=2, notch=TRUE) + 
  theme(axis.text.x = element_text(angle = 30, vjust = 1, size = 10, hjust = 1))
```
It can be seen from the boxplot that leagues which have especially bad players (overall abillity is under 50) are england, france, italy, poland, scotland and "other". Leagues with especially good players (overall ability is over 90) are endland, france and spain. In addition, on average the league with the best players overall abillity is spain. 

```{r}
ggplot(fifa_players_info, aes(x=Wage)) + geom_density()
ggplot(fifa_players_info, aes(x=log(Wage))) + geom_density()
```
note that the log plot does not include 219 players who don't get paid at all (wage=0, log(wage)=-inf) 

```{r}
top_players <- fifa_players_info[, c(2,6,10)]
top_value <- top_players[order(top_players$Value, decreasing = TRUE),]
top_value <- head(top_value,10)
top_value 
top_overall <- top_players[order(top_players$Overall, decreasing = TRUE),]
top_overall <- head(top_overall,10)
top_overall
identical(top_overall, top_value)
  for (i in 1:10){
    if(top_overall[i,1] %in% top_value$Name){ }
    else{
      print(top_overall[i,1])
      break()
    }
  }
```
Two lists are not identical. 
M. Neuer is the player with the best overall ability who is also not in the top 10 valued players. 

```{r}
teams <- aggregate(Overall~Club, fifa_players_info,FUN = mean)
teams <- teams[order(teams$Overall, decreasing = TRUE),]
head<-head(teams, 10)
tail<-tail(teams, 10) 
teams <- cbind(head,tail) 
teams 
```


```{r}
fifa_n <- fifa
fifa_n$Nationality <- as.character(fifa_n$Nationality)
fifa_n$Nationality[fifa_n$Nationality == "England"] <- "UK"
fifa_n$Nationality[fifa_n$Nationality == "Wales"] <- "UK"
fifa_n$Nationality[fifa_n$Nationality == "Scotland"] <- "UK"
fifa_n$Nationality[fifa_n$Nationality == "Republic of Ireland"] <- "UK"
sum_nation <- count(fifa_n, Nationality) #count table
colnames(sum_nation)[colnames(sum_nation)=="n"] <- "Number of players"
map_w_players <- joinCountryData2Map(sum_nation, nameJoinColumn = "Nationality", joinCode = "NAME") #add data to map
mapCountryData (map_w_players, nameColumnToPlot = "Number of players", catMethod = "pretty", colourPalette = RColorBrewer::brewer.pal(7,"Greens"),addLegend = TRUE , aspect = 1)
#create the map
```

```{r}
overall_Nation <- aggregate(Overall ~ Nationality, data = fifa_n, FUN = mean)
colnames(overall_Nation)[colnames(overall_Nation)=="Overall"] <- "average overall quality"
player_maps <- joinCountryData2Map(overall_Nation, nameJoinColumn = "Nationality", joinCode = "NAME")
colourPalette = RColorBrewer::brewer.pal(7,"OrRd")
mapCountryData (player_maps, nameColumnToPlot = "average overall quality", catMethod = "fixedWidth", colourPalette = colourPalette , numCats = 100)
#we would recommend to travel to Algeria, it looks like they dont have lots of players, but they have high overall average quality 
```
```{r}
Overall_Age <- aggregate(Overall ~ Age, data = fifa, FUN = mean)
Overall_Age <- head(Overall_Age, 20)
ggplot(Overall_Age, aes(x=Age, y=Overall)) + 
  geom_point(stat = "identity" ,shape = 5,colour = "dark green", fill = "white", size = 1.5, stroke = 1.5)
```
```{r}
Potential_Age <- aggregate(Potential ~ Age, data = fifa, FUN = mean)
Potential_Age <- head(Potential_Age, 20)
Potential_Age_overall <- full_join(Potential_Age, Overall_Age, by = "Age")
Potential_Age_overall$average_difference <- Potential_Age$Potential - Overall_Age$Overall
ggplot(Potential_Age_overall, aes(x=Age, y=average_difference, )) +
  geom_point(stat = "identity" ,shape = 5,colour = "dark green", fill = "white", size = 1.5, stroke = 1.5)
# In order to find players for future development we should look at the ages with the middle diffrence in the graph (around 20), so they have the potential and with enough overall ability, so in the long run the players will get much better then they already are at that age.
```
```{r}
young <- fifa
young  <- young[which(young$Age <= 21), ]
young  <- young[which(young$Overall > 70), ]
ggplot(young, aes(x=Wage, y=Potential)) +  geom_point(shape = 21, colour = "black", fill = "white", size = 1, stroke = 1)
lm <- lm(Wage ~ Potential, data = young)
ggplot(young, aes(x=Potential, y=Wage)) + 
    geom_point(color = "black") + geom_smooth(color="darkorange", se=FALSE)
young$expected_wage <- -271577.0 + 3607.3*young$Potential
young$diff <-young$expected_wage - young$Wage
young <- young[order(young$diff, decreasing = TRUE), ]
young <- young[1:10,-(12:72)]
young
```


```{r}
players_skills <- fifa_players_attribures[,c(2:12,18:35)]
skills_names <- c(names(players_skills))
corr <- cor(players_skills) 
ggplot(data = melt(corr, id.vars= skills_names), aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile(color = "white") + 
  ggtitle("Correlation between Skills") + 
  scale_fill_gradient2(high = "#0072B2", low = "#D55E00", midpoint = 0, limit = c(-1,1), space = "Lab", name="Correlation") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 10, hjust = 1))
```
Jumping & streanght seem to be least correlated with other skills. 

```{r} 
players_skills <- fifa_players[c(colnames(players_skills),"CAM","CB","CM","RB","RW","ST","Overall")] 
corr <- as.data.frame(cor(players_skills[,1:29],players_skills[,30:36]))
corrplot(t(corr), "square", tl.col = "darkblue", title = "Correlation of key positions & Overall VS Skills") 
Overall_top_corr <- mutate(corr, skills = c(names(players_skills[,1:29])))
Overall_top_corr <- Overall_top_corr[order(Overall_top_corr$Overall, decreasing = TRUE),]
Overall_top_corr <- Overall_top_corr[,c(7,8)]
head(Overall_top_corr,7) 
```

Build your team by selecting six *different* players, one in each of the above positions in such a way that the players chosen are the best in their individual position. If the same player is the best at multiple positions, try to build the team in such a way that maximize the team's overall average score. 
Using the function `radarchart::chartJSRadar`,graph their abilities (individually for all 6 player) in a spider (radarchart) graph. See below an example for such a chart. What is your team's overall average score?
[radar chart](https://en.wikipedia.org/wiki/Radar_chart) (also called 'spider chart')  

```{r}
team <- c(1:6) 
names(team) <- c("CAM","CB","CM","RB","RW","ST") 
team <- as.data.frame(team) 
for (position in c("CAM","CB","CM","RB","RW","ST")){
  fifa_players_positions <-  fifa_players_positions[order(fifa_players_positions[,position], decreasing = TRUE),] 
  team[position,1:6] = fifa_players_positions$ID[1:6]}
team <- t(team) 
# we recived that the same player is the best at both RW and ST. 
# second best at those position is a player who is already taken by other position. 
# therd best player have same overall abillity. 
# therd best player in RW has a better score at it then therd best player in ST has in ST, therefor he was chosen for position RW. 
team[1,5] <- team[3,5]
team <- team[1,] 
team <- data.frame(ID = as.integer(team))
team <- team %>% mutate("Position" = c("CAM","CB","CM","RB","RW","ST"))
players_team <- left_join(team,fifa_players, by = "ID")
rownames(players_team) <- players_team$Name
#our best team is : L. Messi in CAM, Sergio Ramos in CB, T. Kroos in CM, Marcelo in RB, Neymar in RW & Cristiano Ronaldo in ST. WINNING TEAM!
spidar <- players_team[,c(7, 13:23, 29:46)]
spidar <- as.data.frame(t(as.matrix(spidar)))
chartJSRadar(spidar, labs = as.character(row.names(spidar)), title = "Skills of best's team Players") 
```

```{r}
age <- fifa_players[,-c(1:2, 4:11, 23:27, 46:72)]
age <- aggregate(. ~ Age, data = age, FUN = mean)
age <- head(age, 20)
age <- melt(age, id = "Age")
ggplot(age, aes(x = Age, y =value)) +
  geom_line() + 
  xlim(16,35) + 
  ggtitle("Players Abilities by Age") +
  facet_wrap(~variable)
```

```{r}
variables <- 
  fifa_players$Acceleration +
  fifa_players$Aggression +
  fifa_players$Agility +
  fifa_players$Balance +
  fifa_players$Ball.control +
  fifa_players$Composure +
  fifa_players$Crossing +
  fifa_players$Curve +
  fifa_players$Dribbling +
  fifa_players$Finishing +
  fifa_players$Free.kick.accuracy +
  fifa_players$Heading.accuracy +
  fifa_players$Interceptions +
  fifa_players$Jumping +
  fifa_players$Long.passing +
  fifa_players$Long.shots +
  fifa_players$Marking +
  fifa_players$Penalties +
  fifa_players$Positioning +
  fifa_players$Reactions +
  fifa_players$Short.passing +
  fifa_players$Shot.power +
  fifa_players$Sliding.tackle +
  fifa_players$Sprint.speed +
  fifa_players$Stamina +
  fifa_players$Standing.tackle +
  fifa_players$Strength +
  fifa_players$Vision +
  fifa_players$Volleys
levels <- fifa_players
model <- lm(Overall ~ variables, data=levels)
levels$residuals <- model$residuals
levels <- order(levels$residuals)
under_performing <- fifa_players[levels[1:10],]
as_tibble(under_performing)
```



```{r}
# A measure of category's diversity
DIV <- function(category_vec){
  t <- table(category_vec)
  p <- t/sum(t)
  return(sum(p^2))
}
cleaned_data <- fifa_players %>% select(Nationality,Club) %>% na.omit()
number_of_nationality_in_club <- cleaned_data %>% group_by(Club,Nationality) %>% summarise(count = n()) %>% group_by(Club) %>% summarise(N_nation=n()) %>% arrange(desc(N_nation)) %>% mutate(Club = factor(Club,level=unique(Club)))
DIV_in_club <- cleaned_data %>% group_by(Club) %>% summarise(DIV = DIV(Nationality))%>% arrange(DIV)%>% mutate(Club = factor(Club,level=unique(Club)))  # arrange(desc(DIV)) %>% 
# Plot number of different nationalities in each club
g <- ggplot(data = number_of_nationality_in_club %>% head(10),aes(x = Club,y = N_nation,color = Club))
g + geom_bar(stat="identity")
# Plot DIV (diversity?) of different nationalities in each club
g <- ggplot(data = DIV_in_club %>% head(10),aes(x = Club,y = DIV, color = Club))
g <- g + geom_bar(stat="identity")
g
```
In the first plot our predecessor tried to show the top 10 clubs with the highest number of different nationalities. The closest the club to 1 the least div in their club. we can see that the club with the smallest value is the most diverse club.

In the second plot our predecessor tried to show football clubs with the highest number of players' nationalities in their team, according to DIV index and plotted the`10 most diverse  clubs according to this DIV measurement.
Q4.2 :
Q4.2.1 : 
The first problem is that the first graph is not intuitive, because we expect that the higher the DIV value, the more divese a club should be. Here, it's the opposite.
Q4.2.2: 
The second problem in the graphs are the lack of Headlines.
Healines should be included when presenting a plot.
Q4.2.3 : 
The third problem is that people without any club were categorized as people with a club and that all the people without a club were actually categorized in the graphs as one club.That is a problem because these players are not supposed to be included in the diversity of club considurations.
Q4.2.4 : 
The fourth problem is that the graph is hard to read, the names of the clubs on the graph. this is why we thought to changed the angle and the size of the text on the x axis.
Q4.3 : 
```{r}
#changed to show the diver in the right way
DIV <- function(category_vec){
  t <- table(category_vec)
  p <- t/sum(t)
  return(1-sum(p^2))}
#fix the prob of players without a club.
cleaned_data <- fifa_players %>% select(Nationality,Club)
cleaned_data<-cleaned_data[!(cleaned_data$Club==""),]
number_of_nationality_in_club <- cleaned_data %>% group_by(Club,Nationality) %>% summarise(count = n()) %>% group_by(Club) %>% summarise(N_nation=n()) %>% arrange(desc(N_nation)) %>% mutate(Club = factor(Club,level=unique(Club)))
DIV_in_club <- cleaned_data %>% group_by(Club) %>% summarise(DIV = DIV(Nationality))%>% arrange(DIV)%>% mutate(Club = factor(Club,level=unique(Club)))  
# arrange(desc(DIV)) %>% 
# Plot number of different nationalities in each club
g <- ggplot(data = number_of_nationality_in_club %>% head(15),aes(x = Club,y = N_nation,color = Club, fill = Club))
g + geom_bar(stat="identity") + labs(title= "Top10 clubs with with number of diffrent nationalities they have") + ylab("n. of different nationalities") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
# removing players without a club
DIV_in_club <- DIV_in_club[-1,]
# Plot DIV (diversity?) of different nationalities in each club
g <- ggplot(data = DIV_in_club %>% tail(10),aes(x = Club,y = DIV, color = Club ,fill = Club))
g <- g + geom_bar(stat="identity")+  theme(axis.text.x = element_text(size = rel(0.8), angle = 45))+
ggtitle("Top10 clubs with the highest diversity of diffrent nationalities") + ylab("Diversity precent PER club")
g
```


```{r}
fifa_ranking <- data.frame(read.csv(url("https://raw.githubusercontent.com/DataScienceHU/DataAnalysisR_2020/master/data/fifa_ranking.csv")))
fifa_ranking$rank_date <- as.Date(fifa_ranking$rank_date)
#func to get the mean of the overall of thr best 17 players of each country
calculate_mean <- function(df, countries){
  for (i in 1:length(countries)){
  sorted <- fifa_players %>% filter(Nationality == countries[i])
  sorted <- sorted[order(-sorted$Overall),][1:17,]
  df[i,2] <- mean(sorted$Overall)}
  return(df)}
#take all the 2018 data
fifa_ranking <- fifa_ranking %>% filter(rank_date >= "2018-01-18")
#correct names
fifa_ranking$Name <- as.character(fifa_ranking$country_full)
fifa_ranking$Name[68] <- "Ivory Coast"
fifa_ranking$Name[fifa_ranking$Name == "Cape Verde Islands"] <- "Cape Verde"
fifa_ranking$Name[fifa_ranking$Name == "USA"] <- "United States"
fifa_ranking$Name[fifa_ranking$Name == "IR Iran"] <- "Iran"
fifa_ranking$Name[fifa_ranking$Name == "Kyrgyz Republic"] <- "Kyrgyzstan"
#calculate the mean of rank 2018 per country
fifa_ranking <- aggregate(total_points~country_full, fifa_ranking,FUN = mean)
colnames(fifa_ranking)<- c('Name', 'total_points')
#work on fifa 18 data- choose all the countries with 17 players and more
sum <- count(fifa_players, Nationality)
sum <- sum %>% filter(n>= 17)
#nations vec
Nation_vec <- sum$Nationality
#base for result
res <- data.frame(Name = sum$Nationality, Overall = 0)
#use the func and order by overall
res <- calculate_mean(res, Nation_vec)
res <- res[order(-res$Overall),]
#merge the two ds and order by rank of total points
res <- merge(res, fifa_ranking, all = FALSE ,by = c('Name'))
res <- res[order(res$total_points, decreasing = TRUE), ]
#first plot- show the reallife data by rank of the countries- in this order of countries.
barplot(res$total_points[1:25], main = "order of points by the ranking & country", names.arg = res$Name[1:25], xlab = "order of country", ylab = "points-rank", col = colors)
legend("topright", legend = res$Name[1:25], cex=0.6)
#sec plot- see the diff and simulariti between the two plots - the ranking of the countries by the overall mean. 
barplot(res$Overall[1:25], main = "order of points by the fifa18 data", names.arg = res$Name[1:25], xlab = "country", ylab = "overall-rank", col = colors)
legend("topright", legend = res$Name[1:25], cex=0.6)
#You can see that in both graphs the index goes down and this also shows that the relative score tables are correlated.
#see the cor of the two of the two values, so that it can be seen that the variables explain one another and that the correlation is quite high. We have assumed that each team of state consists of at least 17 players, and will select the best players. So we averaged players' capabilities per country, and this measure seems to explain the score in 2018 by 0.77
corr_ <- cor(res[c("Overall","total_points")])
corr_
```
